---
title: 'Homework Five: Functions'
author: "April Wright"
date: "2022-09-23"
output: html_document
---
## Due date: Sept 30, 5 PM.

This homework concerns functions.

1. Make a function that reads in data and does some sort of manipulation of it. Undergrads, use the surveys dataset. Graduates, use your own data. Any manipulation is fine: selection, filtering, value removal. 

```{r}

  hf_plus_weight<-function(data,column8,column9){
        plus<-na.omit(data) %>% 
              mutate(add={{column8}}+{{column9}}) %>% 
              select({{column8}},{{column9}},add)
        return(plus)
            }
     hf_plus_weight(surveys_hm_5,hindfoot_length,weight)
```

2. Add a test to this function. How will you know if it worked, and alert a user if it did not?

```{r}
 hf_plus_weight<-function(data,column8,column9){
            plus<-na.omit(data) %>% 
              mutate(add = {{column8}} + {{column9}}) %>% 
              select({{column8}},{{column9}},add) %>% 
              mutate(add = as.numeric(add))
        return(plus)
 }

hf_plus_weight(surveys_hm_5,hindfoot_length,weight)
  
```

3. Make a function that makes a plot. This can be any type of plot that we covered. Violin, box, scatter. Include the ability to save the plot to a file in the function.

```{r}
hf_vs_weight<-function(data,column8,column9){
     p<-na.omit(data) %>%  
       ggplot(data,mapping=aes(x={{column8}},y={{column9}}))+geom_point()
     ggsave("hfvsweight.pdf")
     return(p)
     
 }
hf_vs_weight(surveys_hm_5,hindfoot_length,weight)
```

4. For saving to a file, users of your function may want the ability to choose their file name. Make the filename an argument so that users can do so.

```{r}
hf_vs_weight<-function(data,column8,column9,filename){
     plot<-na.omit(data) %>% 
       ggplot(data,mapping=aes(x={{column8}},y={{column9}}))+geom_point() 
     ggsave(filename)
     return(plot)
 }
hf_vs_weight(surveys_hm_5,hindfoot_length,weight,"testfile.pdf")

```
