---
title: "Homework One"
output: html_document
---

## Homework One

### Due Aug. 26th at 5pm. Simply write your answers in and save in the homeworks folder in your Rstudio.

This is going to be a low-stakes effort to get you comfortable with RStudio and RMarkdown. First, you're going to download the RMarkdown Document for this homework. Copy and paste this command into the console of your RStudio instance.

```{r}

```

Can you find where the file downloaded to? Look at the help entry for `download.file` if you aren't sure.

Open the file and answer these two questions: 

1. (1pt) What does `download.file` do? 

```
download.file is used download a file from the internet. 
```

2. (1pt) Where did your HW1.Rmd file download to? 

```
The homework 1 file downloaded in the homework folder in the project section of the cloud.
```

Next, we're going to try to execute some code in an RMarkdown document. Press the green triangle.

```{r}
animals <- c("cat", "rat", "dog") #this code concatenated cat, rat, and dog into a vector and was assigned the value, animals. These values in the argument are words, so they are in quotes.
weights <- c(15, 1, 120) #this code concatenated 15, 1, and 120 into a vector and was assigned the value, weights.
animals[c(1,3)] #this asked to concatenate items 1 and 3 from the vector animals which would be cat and dog.
```

What just happened? Annotate the above code with two comments (each worth one point) explaining the code.

We can also embed plots in RMarkdown. The below code should embed a barplot. Not very useful, but let me know if you don't see one show up in your file.

```{r}
barplot(names.arg=animals, height=weights)
```

Did the plot show up OK?

```
Answer yes or no here #yes

```

## Getting to know you all

For the remainder of the assignment, I need to know three things.

1. What type of data will you be working with for your thesis or research project? For example: matrices of phylogenetic characters, spreadsheets of ecological observations, shapefiles from 3D scanning, etc. If you aren't actively doing research, what kinds of scientific questions do you find interesting?
#I find medical microbiolgy interesting. I'm currently taking immunology and really enjoying it.

2. What are you hoping to get out of class?
#I want to get better at organizing data and creating good graphics so that I can use these skills in the job field. I have never coded before this class and I'm hoping to develope coding skills.

3. Are there topics that aren't on the schedule that you'd be interested to see? 
#I can not think of any topics that i would like to cover that are not on the schedule. I am most excited for learning to perform statistical analysis in R.
