---
title: "ProjectOne"
output: html_document
---


# Data Project One

## 100 points
## DUE Sept 15 at 11:59 PM

Below are the questions for the first data practical assignment. This project uses the "FossilAnts.csv" file, located in the data directory for the project. The point value of each question is denoted next to it. A blank cell is below each for your answer; feel free to create more blank cells as needed.

0. (5 pts) Create a directory called `projects`, and in it, a subdirectory called `project_one`. Use `download.files` to get [this file](https://raw.githubusercontent.com/BiologicalDataAnalysis2019/2023/main/vignettes/project_one.Rmd), and [this one](https://raw.githubusercontent.com/BiologicalDataAnalysis2019/2023/main/projects/project_one/FossilAnts.csv) and save them to the `project_one` directory.


1. 5 pts. Import the `tidyverse` package and load the data. The data for this part of the practical is located in the data directory. Save the data in a variable called `project_dat`. Print the data to the screen to ensure it loaded correctly.

```{r}
#install.packages("tidyverse")
library("tidyverse")
project_dat<-FossilAnts
```

2. (5pts) Check the datatypes of each column. There is a column called reference number. This is a static identifier - it should not be changed, and is an indexer used to identify this specimen uniquely. Do we want to treat it as an integer (this is an opinion question - answers may vary).

```
I do not think we should treat them as an integer, the refrence number is a value that you would never take the mean of or want to do calculations with. If we treated them as a character, we could still group by and count them.
```

3. (5 pts) Change the datatype of the reference number column to character. Take a peek at the function as.character().

```{r}
as.character("reference_no")
is.character("reference_no")
```

4. (5 pts) Look at your data. What are the missing data values? In particular, have a look at the `Tribe` column. In your opinion, are these intelligent missing values for the dataset? Why or why not? If not, how would you like to change them? 

```
"None" is not recognized by R to be NA, so these would not be considered an intelligent or readable missing values. A way we can fix this is when we read in FossilAnts.csv from the data directory, we can recognize "None" as NA values. The code would look something like this:
read.csv("FossilAnts.csv",na="None") %>% 
na.omit()
```

6. (5 points) In the surveys dataset, we have genus and species split between two columns. Here, we have them combined. What are the pros and cons of the way we have recorded taxa in these two data sets.

```
Having genus and taxa together could make the data set look less busy when we try to analyse the data, but we have limited ourselves when grouping or counting. We can no longer make groups or count the number of subjects in a specific genus or specific taxa. 
```

7. (5pts) Please look at the help page for the `separate` function.

```{r}
?separate
```


7. (10 pts) How could you separate one column into two? 

```{r}
library(tidyr)
library(dplyr)
FossilAnts %>% 
separate(Ftaxon,c('genus','taxa'))
```

Next, we will test a hypothesis. Your hypothesis is that there are more specimens in the 75 million years ago (mya) - 100 mya interval than the 30 mya to the present interval.

8. (5 pts) Write out the steps you would take to address this question. Will you need to split up the data? Will you need to group the data based on the values in some column? 
```
#First, we could make a data set with four filters. One filter can select orgaisms whose min_ma values are less than 100 million years. Another filter can select organism whose min_ma values are over 75 million years. We can then have another two filters that will also select for organisms above 75 mya and under 100 mya but according to their max-ma values.This ensures we have no organisms that may be outside this range compeltely or partially. After R filters for the orgainsms we want, we can run a count to get the total number of organisms that fit our criteria.

We can then make another data set and code another four filters using the same groups of min_ma and max_ma but the vlaues 30 and 0 that will replace the values of 100 and 75. After the filters are made we can also run a count on organisms that fit our criteria. 

Finally we can combine the two counts into one data table.
```
9. (15 pts) Perform the operations you described in (8). 

```{r}
FossilAnts_75_100<-FossilAnts %>% 
  filter(min_ma<=100) %>% 
  filter(min_ma>=75) %>% 
  filter(max_ma<=100) %>% 
  filter(max_ma>=75) %>% 
  select(reference_no,min_ma,max_ma) %>% 
  count()
FossilAnts_30_0<-FossilAnts %>% 
  filter(min_ma<=30) %>% 
  filter(min_ma>=0) %>% 
  filter(max_ma<=30) %>% 
  filter(max_ma>=0) %>% 
  select(reference_no,min_ma,max_ma) %>% 
  count()
rbind(FossilAnts_75_100,FossilAnts_30_0)

#AMW: Nice work. The fact that there is uncertainty in how old the specimens are makes this slippery.
```

10. (5 pts) Do the results of your code support the hypothesis? 

```
No, the hypothesis is not supported. We only see 48 organism in the 100 mya to 75 mya. We see 313 organisms in the 30 mya to 0 mya range. Since there are more organisms counted in the more present region, the evidence shows that there are less orgaisms in the 100-75 mya range than the 30-0 mya.
```

11. (10 pts) Save the dataframe with the split taxon columns into a new directory called `project_one_data_output`. Save it as a csv file called "column_separated.csv" 

```{r}
FossilAnts_Q11 <- read_csv("project_one_data_output/FossilAnts_Q11.csv")
#AMW: Fixed path.
```

12. (10 pts) Produce a histogram of the number of specimens by minimum age. Look at the geom_histogram() help.

```{r}
ggplot(FossilAnts_Q11,aes(min_ma))+geom_histogram(binwidth=2)

```

13. (5 pts) Change the font size on your histogram so that it can be read comfortably on your computer screen from five feet away. Save the file as "large_font.pdf".

```{r}
ggplot(FossilAnts_Q11,aes(min_ma))+geom_histogram(binwidth=2)+theme(text=element_text(size=24))
ggsave("large_font.pdf")
```

14. Finally, produce a histogram of counts for each subfamily.

```{r}
ggplot(FossilAnts_Q11,aes(min_ma))+geom_histogram(binwidth=2)+facet_wrap(vars(subfamily))

```


# Graduate Students

1. You will do the above steps with the classroom dataset.
2. Next, you will choose three data steps from above to perform with your own data. This could be grouping, mutating, cleaning NA values, visualization, or any combination of the three. 
3. Why did you choose these steps? What function does this serve for your thesis and/or other research projects? 

